---
title: "PrelimProjectLookGroupC"
author: "Sam Schneider"
date: "4/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Read in the data**

```{r}
require(readr)
require(mosaic)
groupcdata <- read_csv("https://awagaman.people.amherst.edu/stat230/projectsS18/groupCdataS18.csv")

```

**Summary command on the data set**

## Put data in right format.
```{r}
summary(groupcdata)
#Change Year, Month and Date to right format
sales<-groupcdata %>% 
  mutate(condition=as.factor(condition),id=1:n(),Year=as.factor(parse_number(Year)),Month=as.factor(parse_number(Month)),Date=as.factor(parse_number(Date)),waterfront=as.factor(waterfront),zipcode=as.factor(zipcode)) %>% 
  rename(Day=Date) 


```

**Data Codebook**


Our variables are: 

Bedrooms - The number of bedrooms in the house (numeric)   
Bathrooms - The number of bathrooms in the house (numeric)   
sqft_lot - The total square footage of the property (numeric)  
floors - The number of floors in the house (numeric)  
waterfront - Indicator if the house has a waterfront view (categorical)  
condition - The overall condition of the house based upon the King County grading system (categorical)  
yr_built - The year in which the house was built (numeric)
zipcode = The ZIP code in which the house is located (categorical)

**Analysis Plan**  

1.	How much does the location of the home determine the value of that home? (Response Variable = Price , Explanatory Variable(s)= ZIP Code)
2.	How much does having a waterfront view affect the impact of home location in determining home value?(Response Variable = Price , Explanatory Variable(s)= ZIP Code, Waterfront View, interaction term - ZIP Code : Waterfront View)
3.	Is the square footage of a house a better predictor of home value than the number of bedrooms?(Response Variable = Price, Explanatory Variable(s)= Square Footage, Bedrooms)

**Prelim Univariate Analysis**
```{r}

```
```{r}
fav_stats(groupcdata$price)
densityplot(groupcdata$price)
```

The average price of a home in this market is $540,000 although this number is likely subject to outliers as shown by the density plot. Accordingly, the median is a lower value of $450,000.

```{r}
histogram(groupcdata$condition)
histogram(groupcdata$floors)

```
The majority of houses in this market only have one floor and it is very rare to have more than 2 floors. Furthermore, the majority of the homes are in above average condition as most homes are graded at a 3 or above. 

```{r}
favstats(groupcdata$sqft_lot)
densityplot(groupcdata$sqft_lot)
```
The values for sqft in this market is extremely distorted by outliers. The mean value is over 15,000 square feet while the median is roughly half of the mean. This will be important to consider as we continue our analysis.

```{r}
tally(groupcdata$waterfront)
```
It is very rare to have a waterfront home in this market. This could make it very valuable in terms of the price of the home.

**Prelim Bivariate Analysis**

**Randomization-Bsaed Procedure Thoughts**
