---
title: "PrelimProjectLookGroupC"
author: "Sam Schneider, Raymond Saitoti, Brian Burrows"
date: "4/9/2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small", warning=FALSE, message=FALSE)
```

**Read in the data**

```{r}
require(readr)
require(mosaic)
groupcdata <- read_csv("https://awagaman.people.amherst.edu/stat230/projectsS18/groupCdataS18.csv")

```

**Summary command on the data set**

## Put data in right format.
```{r}
summary(groupcdata)
#Change Year, Month and Date to right format
#Filter zipcodes of interest
sales<-groupcdata %>% 
  mutate(condition=as.factor(condition),id=1:n(),waterfront=as.factor(waterfront),zipcode=as.factor(zipcode)) %>% 
  filter(zipcode %in% c("98103","98038","98115","98052","98117"))
```

**Data Codebook**


Our variables are: 

bedrooms - The number of bedrooms in the house (numeric: 0-33)   
bathrooms - The number of bathrooms in the house (numeric: 0-8)   
sqft_lot - The total square footage of the property (numeric: 520-1651359)  
floors - The number of floors in the house (numeric: 1-3.5)  
waterfront - Indicator if the house has a waterfront view (categorical: 0-1, where 0 is no waterfront view)  
condition - The overall condition of the house based upon the King County grading system (categorical: 0-5)  
yr_built - The year in which the house was built (numeric: 5/2/14 - 5/27/15)
zipcode - The ZIP code in which the house is located (categorical)
price - The price a house was sold for (numeric: 75000-7700000)
sqft_living - The total square footage of the house (numeric: 370-5770)


**Analysis Plan**  

As stated in the proposal, we aim to create a model that can predict home values in King County, using 2014 and 2015 sales data. In addition, we will examine the following additional questions:

1. How much does the location of the home determine the value of that home? (Response Variable = Price , Explanatory Variable(s)= ZIP Code)
2. How much does having a waterfront view affect the impact of home location in determining home value?(Response Variable = Price , Explanatory Variable(s)= ZIP Code, Waterfront View, interaction term - ZIP Code : Waterfront View)
3. Is the square footage of a house a better predictor of home value than the number of bedrooms?(Response Variable = Price, Explanatory Variable(s)= Square Footage, Bedrooms)

**Prelim Univariate Analysis**

```{r}
fav_stats(sales$price)
densityplot(sales$price)
```

The average price of a home in this market is $557,996 although this number is likely subject to outliers as shown by the density plot. Accordingly, the median is a lower value of $525,000.

```{r}
histogram(sales$condition)
histogram(sales$floors)
```

The majority of houses in this market only have one floor and it is very rare to have more than 2 floors. Furthermore, the majority of the homes are in above average condition as most homes are graded at a 3 or above. 

```{r}
favstats(sales$sqft_lot)
densityplot(sales$sqft_lot)
```

The values for sqft in this market is extremely distorted by outliers. The mean value is roughly 10,000 square feet while the median is roughly half of the mean. This will be an important factor to consider as we continue our analysis.

```{r}
tally(sales$waterfront)
```
It is very rare to have a waterfront home in this market. This could make it very valuable in terms of the price of the home.

**Prelim Bivariate Analysis**

We begin with simple analyses of the relationships price has with each of our predictors to see if transformations are necessary.


First is price and bedrooms.

```{r}
xyplot(price~bedrooms, main="House Price vs. Number of Bedrooms", data=sales)
with(sales, cor(price,bedrooms))
```
From the graph, we see that baseline home values seem to increase as the the number of bedrooms increases. However, home values seem to increase up to six bedrooms and decrease afterwards. This might most likely be the result of a small sample size for houses with more bedrooms than that. It might be worthwhile to consider bedrooms and zipcode together in future analysis. There is also one extreme outlier with more than thirty rooms. We might want to consider looking more closely at that point to determine if we should remove it. Also, as expected, home price and number of bedrooms seems to be positively correlated, although weakly so. 

Second is price and bathrooms.

```{r}
xyplot(price~bathrooms, main="House Price vs. Number of Bathrooms", data=sales)
with(sales, cor(price,bathrooms))
```

The graph indicates a moderate positive relationship between price and number of bathrooms. However, it is worthwhile noting that there are far fewer observations for houses with five or more bathrooms. This might limit the scope to which we may predictions out of fear of extrapolation beyond the data.

Third is price and lot square footage.

```{r}
xyplot(price~sqft_lot, main="House Price vs. Square Footage", data=sales)
with(sales, cor(price,sqft_lot))
```

The graph of house price and square footage indicates that transformations will most likely be required. Furthermore, very few observations exist for homes of more than 500,000 square feet. In addition, a correlation test indicates that price and square footage is very weakly positively correlated, which is expectated but not readily shown by the graph. 

```{r}
xyplot(price~log(sqft_lot), main="House Price vs. Square Footage", data=sales)
#This transformation that I think looks best. However, the data does not exhibit much of a linear relationship and might not be worth using in the final model. 
with(sales, cor(price,log(sqft_lot)))
```

A log transformation of lot square footage improves the legibility of the graph and strengthens the correlation between square footage and house price.

Fourth is price and floors.

```{r}
xyplot(price~floors, main="House Price vs. Number of Floors", data=sales)
with(sales, cor(price,floors))
```

While the scatterplot was not particularly informative, a correlation test shows that price and number of floors are weakly positively correlated. However, the scatterplot does elucidate a number of high outliers that should be examined more closely, especially the one for 2.5 floors.

Fifth is price and waterfront.

```{r}
bwplot(price~waterfront, main="House Price vs Waterfront View", xlab="Waterfront View", data=sales)
```

Side by side boxplots of house price by waterfront view supports earlier univariate analysis as waterfront properties seem to have higher median price. Whether this difference is significant is something to be evaluated in a regression.

Sixth is price and condition.

```{r}
bwplot(price~condition, main="House Price vs Home Condition", xlab=" Home Condition", data=sales)
```

The side-by-side boxplots of price by house condition shows fairly similar medians for each home condition group. Interestingly, there are a number of outliers for homes with conditions rated three or higher. This might be explained by location or age data in additional analysis. Also, an ANOVA will be run to see if the differences in house price between condition groups is significant.


Seventh is price and zipcode.

```{r}
bwplot(price~zipcode, main="House Price vs Zipcode", data=sales)
```


Boxplots of price by zip-code seem to indicate differences in house price by zip-code. An ANOVA will be run to see if any of the differences are significant. It also seems like there are alot of outliers for each of the zipcodes.

Last is price vs. house/living square footage

```{r}
xyplot(price~sqft_living, data=sales)
with(sales, cor(price,sqft_living))
```

The scatterplot of price and house square footage shows a weak to moderate positive linear correlation. The correlation coefficient is approximately 0.61, which is moderately strong. 

## Spatial analysis of house concentration in the five most popular zipcodes:

```{r}
library(ggmap)
library(mapproj)
#Heatmap showing where the sales are happening across all ZIP codes, use this to show why we are focusing on the 5 zip codes
map <- get_map(location = 'Seattle')
ggmap(map, extent = "device") + geom_density2d(data = groupcdata, aes(x = long, y = lat)) + stat_density2d(data = groupcdata, aes(x = long, y = lat, fill = ..level.., alpha = ..level..), size = .01, bins = 16, geom = "polygon") + scale_fill_gradient(low = "yellow", high = "red") +scale_alpha_continuous(guide = FALSE) + ggtitle("Where Sales Are Happening in King County")
```

```{r}
ggmap(map) +
  geom_point(data = sales, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_size(range = c(3, 8)) + ggtitle("ZIP Codes with Most Sales in King County") + scale_alpha_continuous()
```



Sale Prices in 98038 ZIP Code
```{r}

map <- get_map(location = '98038', zoom = 12)

sales_98038 <- sales %>%
  filter(zipcode == "98038")
ggmap(map) +
  geom_point(data = sales_98038, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_color_manual(values = c("darkgreen"), guide = FALSE) + ggtitle("Distribution of Sale Price in 98038 ZIP code") + scale_alpha_continuous()
```

Sale Prices in 98103 ZIP Code
```{r}
map <- get_map(location = '98103', zoom = 13)

sales_98103 <- sales %>%
  filter(zipcode == "98103")
ggmap(map) +
  geom_point(data = sales_98103, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_color_manual(values = c("darkgreen"), guide = FALSE) + ggtitle("Distribution of Sale Price in 98103 ZIP code") + scale_alpha_continuous()
```

Sale Prices in 98052 ZIP Code
```{r}
map <- get_map(location = '98052', zoom = 11)

sales_98052 <- sales %>%
  filter(zipcode == "98052")
ggmap(map) +
  geom_point(data = sales_98052, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_color_manual(values = c("darkgreen"), guide = FALSE) + ggtitle("Distribution of Sale Price in 98052 ZIP code") + scale_alpha_continuous()
```

Sale Prices in 98115 ZIP Code
```{r}
map <- get_map(location = '98115', zoom = 13)

sales_98115 <- sales %>%
  filter(zipcode == "98115")
ggmap(map) +
  geom_point(data = sales_98115, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_color_manual(values = c("darkgreen"), guide = FALSE) + ggtitle("Distribution of Sale Price in 98115 ZIP code") + scale_alpha_continuous()
```

Sale Prices in 98117 ZIP Code
```{r}
map <- get_map(location = '98117', zoom = 14)

sales_98117 <- sales %>%
  filter(zipcode == "98117")
ggmap(map) +
  geom_point(data = sales_98117, aes(x = long, y = lat, color = zipcode, alpha = price)) +
  labs(x="Longitude", y="Latitude")+ 
    scale_color_manual(values = c("darkgreen"), guide = FALSE) + ggtitle("Distribution of Sale Price in 98117 ZIP code") + scale_alpha_continuous()
```

## Models

## **Randomization-Based Procedure**

We suspect that our price vs square footage plot might need a log transformation. However, we would like to see whether the results we see after fixing the conditions would be compatible with those obtained using a randomization test for the slope. The variable we would be shuffling in this context would be the square footage

```{r}
xyplot(log(price)~sqft_lot, main="House Price vs. Square Footage", data=sales)
with(sales, cor(price,sqft_lot))
fm<-lm(log(price)~sqft_lot,data=sales)
plot(fm,which=1)
plot(fm,which=2)
```

After doing a log transformation on price, it is clear that some of the conditions for the linear model still haven't been met. There is still a problem with the distribution of error terms as seen in the variation across the fitted values. Thus, we shall run the randomization test below to test the significance of the relationship between price and square footage.

```{r}
t<-do(1000)*(cor(shuffle(sales$sqft_lot),sales$price))
hist(t$result)
with(sales, cor(price,sqft_lot))
savecor<-t$result
pdata(savecor,.036,lower.tail=FALSE) 
qdata(savecor, c(0.025, 0.975))
```

## Anova analysis to compare how much the location of a home determined its value

### One way.
```{r}
modl<-lm(log(price)~zipcode,data=sales)
anova(modl)
plot(modl,which=1)
plot(modl,which=2)
```

Conditions:
There are a few outliers on the upper and lower tails of the qqplot. Constant variance is satisfied as can be seen in the residuals vs fitted plot

the F-statistic of 309.8 and its associated p-value of 2.2e-16 provide evidence that there is at least one difference between the zipcodes in terms of average price. 

**Pairwise comparisons showing where differences lie **

```{r}
#Tukeys LSD
require(DescTools)
PostHocTest(aov(log(price)~zipcode,data=sales), method = "hsd")
```

### Two way ANOVA

```{r}
#Two way modes + interaction
sales<-sales %>% 
  mutate(zipcode=factor(zipcode))
modl2<-lm(log(price)~zipcode+waterfront,data=sales)
anova(modl)
plot(modl2,which=1)
plot(modl2,which=2)
```

## Stepwise Regression Model

```{r}
modall <- lm(log(price)~bedrooms+bathrooms+sqft_lot+log(sqft_lot)+floors+waterfront+condition+yr_built+zipcode+sqft_living, data=sales)
summary(modall)
plot(modall,which=1)
plot(modall,which=2)
```

## Results

**Results**
The observed correlation between square footage and price is 0.036. After using a randomization test to shuffle up the square footage, 1000 correlations are obtained. Their resulting distribution shows that 0.036 falls within the constructed interval after destroying the relationship between price and square footage, which suggests that the correlation between these two variables is not significantly greater than zero (House price is unrelated to square footage). The one way ANOVA ran showed that there existed at least one difference in average house prices between the zipcodes analyzed. The multiple comparisons ran between average house prices in different zipcodes revealed that significant differences exist between every pairwise combination of zipcodes except 98117(Loyal Heights/Ballard neighborhood) and 98103(Fremont/WallingFord neighborhood). The test for interaction between zipcode and waterfront yields a high p value (0.269) which shows that the interaction term is insignificant  after zipcode and waterfront have been accounted for.



