---
title: "PrelimProjectLookGroupC"
author: "Sam Schneider"
date: "4/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Read in the data**

```{r}
require(readr)
groupcdata <- read_csv("https://awagaman.people.amherst.edu/stat230/projectsS18/groupCdataS18.csv")

```

**Summary command on the data set**

## Put data in right format.
```{r}
summary(groupcdata)
#Change Year, Month and Date to right format
sales<-groupcdata %>% 
  mutate(condition=as.factor(condition),id=1:n(),Year=as.factor(parse_number(Year)),Month=as.factor(parse_number(Month)),Date=as.factor(parse_number(Date)),waterfront=as.factor(waterfront),zipcode=as.factor(zipcode)) %>% 
  rename(Day=Date) 


```

**Data Codebook**


Our variables are: 

Bedrooms - The number of bedrooms in the house (numeric)   
Bathrooms - The number of bathrooms in the house (numeric)   
sqft_lot - The total square footage of the property (numeric)  
floors - The number of floors in the house (numeric)  
waterfront - Indicator if the house has a waterfront view (categorical)  
condition - The overall condition of the house based upon the King County grading system (categorical)  
yr_built - The year in which the house was built (numeric)
zipcode = The ZIP code in which the house is located (categorical)

**Analysis Plan**  

1.	How much does the location of the home determine the value of that home? (Response Variable = Price , Explanatory Variable(s)= ZIP Code)
2.	How much does having a waterfront view affect the impact of home location in determining home value?(Response Variable = Price , Explanatory Variable(s)= ZIP Code, Waterfront View, interaction term - ZIP Code : Waterfront View)
3.	Is the square footage of a house a better predictor of home value than the number of bedrooms?(Response Variable = Price, Explanatory Variable(s)= Square Footage, Bedrooms)

**Prelim Univariate Analysis**
```{r}

```
**Prelim Bivariate Analysis**

**Randomization-Bsaed Procedure Thoughts**
